Kids PM Summary

### 2019-12-06 data_cleaning_initial.Rmd
dusttrak data cleaning

# Load, format, and save KidsAIR Dusttrack data
```{r}
##### Read in dusttrack .txt files (from WMT) #####

@@

# Save pm dataset as RDS and CSV
write_rds(kids_pm, "Output/kids_pm_full.rds")
#write_csv(kids_pm, "Output/kids_pm_full.csv")

# Load data and clean up
kids_pm <- read_rds("Output/kids_pm_full.rds")

kids_pm_check <- kids_pm %>% 
  group_by(home_winter_id) %>% 
  summarise(obs = n()) %>% 
  ungroup()

ak_dt_home <- read_csv("Input/AK/ak_dt_home.csv")
nn_dt_home <- read_csv("Input/NN/nn_dt_home.csv")
wmt_dt_home <- read_csv("Input/WMT/wmt_dt_home.csv")

dt_home <- rbind(ak_dt_home, nn_dt_home, wmt_dt_home) %>% 
  rename_all(tolower) %>% 
  filter(!is.na(homewinterid)) %>% 
  mutate(home_winter_id = as.character(homewinterid)) %>% 
  left_join(kids_pm_check, by = "home_winter_id")

##### Add to new variables to Kids PM data #####
kids_pm_new <- kids_pm %>% 
  mutate(home_winter_id = as.character(home_winter_id),
         area = as.factor(area),
         pm_date = date,
         pm_time = time,
         pm_datetime = datetime,
         pm = as.numeric(pm)) %>% 
  select(home_winter_id, area, pm_date, pm_time, pm_datetime, pm) %>% 
  arrange(home_winter_id, pm_date) %>% 
  group_by(home_winter_id) %>% 
  # new variables summarizing pm2.5 within each home_winter_id
  mutate(pm_total_observations = n(),
         pm_total_summed = sum(pm),
         min_datetime = first(pm_datetime),
         max_datetime = last(pm_datetime),
         min_datetime = ymd_hms(min_datetime),
         max_datetime = ymd_hms(max_datetime),
         diff_datetime = interval(min_datetime, max_datetime),
         pm_total_days = diff_datetime/86400,
         pm_24hr_mean = pm_total_summed/pm_total_days,
         pm_rolling_24hr = rollsum(pm, 1440, fill = "extend", align = "right"),
         obs = row_number(),
         sampling_day = if_else(obs < 1441, 1, 0),
         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
  group_by(home_winter_id, sampling_day) %>% 
  mutate(pm_24hr_daily = sum(pm),
         pm_24hr_daily_weighted = pm_24hr_daily*(1440/n()),
         pm_sampling_day = sampling_day) %>% 
  ungroup() %>% 
  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day)

>>>>

# Load, format, and save KidsAIR Dusttrak data
```{r}
##### Read in dusttrack .txt files (from WMT) #####

@@

# Save pm dataset as RDS and CSV
write_rds(kids_pm, "Output/kids_pm_full.rds")
#write_csv(kids_pm, "Output/kids_pm_full.csv")
```

# Load Dusttrak data and clean up
```{r}
kids_pm_full <- read_rds("Output/kids_pm_full.rds")

kids_pm_check <- kids_pm_full %>% 
  group_by(area, home_winter_id) %>% 
  summarise(obs = n()) %>% 
  ungroup()

ak_dt_home <- read_csv("Input/AK/ak_dt_home.csv") %>% 
  mutate(area = "AK")
nn_dt_home <- read_csv("Input/NN/nn_dt_home.csv") %>% 
  mutate(area = "NN")
wmt_dt_home <- read_csv("Input/WMT/wmt_dt_home.csv") %>% 
  mutate(area = "WMT")

dt_home <- rbind(ak_dt_home, nn_dt_home, wmt_dt_home) %>% 
  rename_all(tolower) %>% 
  filter(!is.na(homewinterid)) %>% 
  mutate(home_winter_id = as.character(homewinterid)) %>% 
  left_join(kids_pm_check, by = c("area", "home_winter_id"))

%>% 
  group_by(home_winter_id) %>% 
  mutate(n_hwid = n()) %>% 
  ungroup() %>% 
  filter(n_hwid == 2)

##### Add to new variables to Kids PM data #####
kids_pm_new <- kids_pm_full %>%  
  #filter(area == "AK") %>% 
  # apply correction factor to pm
  mutate(pm = pm/1.65) %>% 
  mutate(home_winter_id = as.character(home_winter_id),
         area = as.factor(area),
         pm_date = date,
         pm_time = time,
         pm = as.numeric(pm),
         day_of_week = weekdays(date)) %>% 
  unite("pm_datetime", date:time, sep = " ") %>% 
  select(home_winter_id, area, pm_date, pm_time, pm_datetime, day_of_week, pm) %>% 
  arrange(area, home_winter_id, pm_date) %>% 
  group_by(area, home_winter_id) %>% 
  # new variables summarizing pm2.5 within each home_winter_id
  mutate(pm_total_observations = n(),
         pm_total_summed = sum(pm),
         # new datetime values to calculate sampling interval
         min_datetime = first(pm_datetime),
         max_datetime = last(pm_datetime),
         min_datetime = ymd_hms(min_datetime),
         max_datetime = ymd_hms(max_datetime),
         diff_datetime = interval(min_datetime, max_datetime),
         # total sampling interval, based on first and last datetimes
         pm_sample_interval = diff_datetime/86400,
         # expected number of observations with the sampling interval
         expected_obs = pm_sample_interval*24*60,
         # difference between actual and expected observations
         actual_vs_expected_diff = pm_total_observations - expected_obs,
         # indicator variable if actual vs expected obs > 10 or greater
         actual_vs_expected = if_else(actual_vs_expected_diff < 10 &
                                      actual_vs_expected_diff > -10, 1, 0),
         # mean pm based on 24hr periods in the sampling interval
         pm_24hr_mean_interval = pm_total_summed/pm_sample_interval,
         # mean pm based on 24hr periods of total observations
         pm_24hr_mean_obs = pm_total_summed/(pm_total_observations/1440),
         # rolling 24hr mean pm
         pm_rolling_24hr = rollsum(pm, 1440, fill = "extend", align = "right"),
         # breaking sampling period into days
         obs = row_number(),
         sampling_day = if_else(obs < 1441, 1, 0),
         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
  # calculate pm means for each sampling day
  group_by(home_winter_id, sampling_day) %>% 
  mutate(pm_24hr_daily = sum(pm),
         pm_24hr_daily_weighted = pm_24hr_daily*(1440/n()),
         pm_sampling_day = sampling_day) %>% 
  ungroup() %>% 
  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day)

### 2019-12-06 pm_summary.Rmd
dusttrak data cleaning

# Summary stats
```{r}
summary_kids_pm <- kids_pm %>% 
  na.exclude() %>% 
  #filter(pm < 0)
  summarize(mean(pm_24hr_mean), sd(pm_24hr_mean),
            min(pm_24hr_mean), max(pm_24hr_mean),
            min(pm), max(pm), mean(pm_total_days),
            min(pm_total_days), max(pm_total_days),
            min(pm_total_observations), max(pm_total_observations),
            mean(pm_24hr_daily), mean(pm_24hr_daily_weighted)) %>% 
  t()
summary_kids_pm
```

>>>>

# Summary stats
```{r}
summary <- kids_pm %>% 
  filter(pm < 0)

summary_kids_pm <- kids_pm %>%
  #filter(pm < 0)
  summarize(mean(pm_24hr_mean_interval, na.rm = TRUE), 
            sd(pm_24hr_mean_interval, na.rm = TRUE),
            min(pm_24hr_mean_interval, na.rm = TRUE), 
            max(pm_24hr_mean_interval, na.rm = TRUE),
            mean(pm_24hr_mean_obs, na.rm = TRUE), 
            sd(pm_24hr_mean_obs, na.rm = TRUE),
            min(pm_24hr_mean_obs, na.rm = TRUE), 
            max(pm_24hr_mean_obs, na.rm = TRUE),
            min(pm, na.rm = TRUE), 
            max(pm, na.rm = TRUE), 
            mean(pm_sample_interval, na.rm = TRUE),
            min(pm_sample_interval, na.rm = TRUE), 
            max(pm_sample_interval, na.rm = TRUE),
            min(pm_total_observations, na.rm = TRUE), 
            max(pm_total_observations, na.rm = TRUE),
            mean(pm_24hr_daily, na.rm = TRUE), 
            mean(pm_24hr_daily_weighted, na.rm = TRUE)) %>% 
  t()
kable(summary_kids_pm, digits = 2)
```
